Conversion of FFCS ICE fleet to EV: a data-driven approach

### 2020-12-04 Chapter1/1_0_context.tex
added appendix

%% Increasing of population in big cities and relative mobility demand
Nowadays, the 55\% world population is strongly concentrated in urban centers and the authors of \cite{UNfuture} forecasted, in 2018, an increase until 68\% before 2050. In this scenario, one of the problem  afflicts big conurbations is mobility. In the past decades, the economic growth were sustain by the flourishing automotive industry. \cite{8_matas2008changes} demonstrated how the presence of private cars in an European Countries like Spain will follow the demographic and correlated mobility demand increase.

%% cost related to traffic congestion (emission related to TC  and pollutant mortality costs)
The increase of private vehicles derived from the increased mobility demand leads to an higher probability of traffic congestion, especially in rush hours. A lot of past and recent studies demonstrated how traffic jams have a negative effect on the public health. Indeed  \cite{12_leiriao2020environmental} showed how the thin dust emission are strongly related with traffic, measuring air pollutant before and during a truck strike. Moreover,  \cite{13_levy2010evaluation} computed the monetized value of PM2.5-related emission attributable to congestion. The results are shocking: the public health may cost up to \$17 billion in 2030 in the United States. Another aspect that decreases the quality of life in congested city is the noise pollution. In particular, \cite{14_mehdi2011spatio} and \cite{15_jacyna2017noise} stated how the noise can reach very dangerous level and causes permanent damages to people living close to trafficked areas. Finally, the land use  related to cars presence is another critical point that policy maker must manage.

In the past years, the several different regulatory boards addressed the congestion increasing and optimizing the road network. According to several works like \cite{10_sweet2011does, 16_hymel2010induced, 17_naess2012traffic} adding more resources, attracts more car, actually feeding the traffic congestion burden.

>>>>

%% Increasing of population in big cities and relative mobility demand
Nowadays, the 55\% world population is strongly concentrated in urban centers and the authors of \cite{UNfuture} forecasted, in 2018, an increase until 68\% before 2050. In this scenario, one of the problem that afflicts big conurbations is mobility. In the past decades, the economic growth were sustain by the flourishing automotive industry. \cite{8_matas2008changes} demonstrated how the presence of private cars in an European Countries like Spain follows the demographic and correlated mobility demand increase.

%% cost related to traffic congestion (emission related to TC  and pollutant mortality costs)
The increase of private vehicles derived from the increased mobility demand leads to an higher probability of traffic congestion, especially in rush hours. A lot of past and recent studies demonstrated how traffic jams have a negative effect on the public health. Indeed  \cite{12_leiriao2020environmental} showed how the micro dust concentration are strongly related with traffic, measuring air pollutant before and during a truck strike. Moreover,  \cite{13_levy2010evaluation} computed the monetized value of PM2.5-related emission attributable to congestion. The results are shocking: the public health may cost up to \$17 billion in 2030 in the United States. Another aspect that decreases the quality of life in congested city is noise pollution. In particular, \cite{14_mehdi2011spatio} and \cite{15_jacyna2017noise} stated how the noise can reach very dangerous level and causes permanent damages to people living close to congested areas. Finally, the land use  related to cars presence is another critical point that policy maker must manage.

In the past years,  several different regulatory boards addressed the congestion increasing and optimizing the road network. According to several works like \cite{10_sweet2011does, 16_hymel2010induced, 17_naess2012traffic} adding more resources, attracts more car, actually feeding the traffic congestion burden.

### 2020-12-04 Chapter11/11_0_conclusion.tex
added appendix

In this thesis, I proposed a data-driven pipeline to study the conversion from internal combustion engine electric vehicles in FFCS. Strongly relying on real FFCS, optimization algorithms, and machine learning techniques I depicted a pipeline able to evaluate the performances of electric FFCS in hot and cold start-up scenarios.

In the first part of my thesis, I described the data gathering, scraping real users' ride from operative combustion engine FFCS. Then I characterized the users' trends in 2017 for two FFCS providers: car2go and Enjoy. Then I highlighted the difference between one-way CS, two-ways CS, and FFCS, which pointed out how the users prefer flexibility over costs. Once all the data are consolidated I developed an electric FFCS mobility simulator. It is an event-based trace-driven simulator. It takes as input the list of rentals events, the infrastructure setups (in terms of charging station placements, car consumptions), and the provider's fleet management policies. It replicates users' patterns and returns as output the metrics describe system performances (complete battery depletion, cars unavailability) and users' discomfort caused car plugging procedures. Armed with this, I initially investigated the best electric FFCS setup varying case of study city (and thus users' patterns), finding how an electric FFCS is sustainable with few charging stations. Next, I improved the solution of the previous step by running several optimization algorithms maximizing both system resilience and minimizing the users' discomfort. Then, I moved my attention to users' ride predictability computing how weather and socio-economics aspects are related to the users' trips. Finally, I simulated possible electric FFCS growth targeting which are the most sensible parameters and how profitable an electric FFCS may be.

@@

However, the research presents several limitations. In particular, the dataset is composed of passive measurements, the reason for that it is impossible to distinguish users' trips and unallocated missed trips. The software does not really know when the users did not find an available vehicle nearby. Moreover, the real traveled distances are unknown, due to the anonymization related to privacy regimentation. This should not really impact macro analyses but it becomes sensible when it necessary to perform neighborhood-grained optimizations. This dissertation does not take into account fleet management (relocation). This topic opens several research threads, that need a pragmatic formulation and sophisticated upgrade to the software. Finally, the recent COVID-19 pandemic changed drastically both private and shared users' patterns opening new questions on mobility sustainment in smart cities.

In general, I believe my thesis proposes a pragmatic methodology to study the (r)evolution of shared mobility into electricity. It started during in course during the Master's degree program, it continued in a Master's Thesis and lasted the doctoral program. It will be very challenging and stimulating for me to face the unsolved problem in my future research career.

>>>>

In this thesis, I proposed a data-driven pipeline to study the conversion from internal combustion engine electric vehicles in FFCS. Strongly relying on real FFCS data, optimization algorithms, and machine learning techniques, I depicted a pipeline able to evaluate the performances of electric FFCS in hot and cold start-up scenarios.

In the first part of my thesis, I described the data gathering, scraping real users' ride from operative combustion engine FFCS. Then I characterized the users' habits in 2017 for two FFCS providers: car2go and Enjoy. I highlighted the difference between one-way CS, two-ways CS, and FFCS, which pointed out how the users prefer flexibility over costs. Once all the data are consolidated I developed an electric FFCS mobility simulator. It is an event-based trace-driven simulator. It takes as input the list of rentals events, the infrastructure setups (in terms of charging station placements, car consumptions), and the provider's fleet management policies. It replicates users' patterns and returns as output the metrics describe system performance (complete battery depletion, cars unavailability) and users' discomfort caused car plugging procedures. Armed with this, I initially investigated the best electric FFCS setup varying case of study city (and thus users' patterns), finding how an electric FFCS is sustainable with few charging stations. Next, I improved the solution of the previous step by running several optimization algorithms maximizing both system resilience and minimizing the users' discomfort. Then, I moved my attention to users' ride predictability computing how weather and socio-economics aspects are related to the users' trips. Finally, I simulated possible electric FFCS growth targeting which are the most sensible parameters and how profitable an electric FFCS may be.

@@

However, the research presents several limitations. In particular, the dataset is composed of passive measurements, the reason for that it is impossible to distinguish users' trips and unallocated missed trips. The software does not really know when the users did not find an available vehicle nearby. Moreover, the real travelled distances are unknown, due to the anonymization related to privacy. This should not really impact macro analyses but it becomes sensible when it necessary to perform neighbourhood-grained optimizations. This dissertation does not take into account fleet management (relocation). This topic opens several research threads, that need a pragmatic formulation and sophisticated upgrade to the software. Finally, the recent COVID-19 pandemic changed drastically both private and shared users' patterns opening new questions on mobility sustainment in smart cities.

In general, I believe my thesis proposes a pragmatic methodology to study the electric (r)evolution of shared mobility. It will be very challenging and stimulating for me to face the unsolved problem in my future research career.

### 2020-12-04 Chapter2/2_3_methodology.tex
added appendix

\section{Data Normalization and Integration}
\label{sec:2_4_data_normalization}
In this second module I illustrates hoe \tool processes and consolidate each snapshot to obtain \emph{parking} and \emph{bookings} periods for each car. A \emph{parking} is time where the car is available for a user ride. On the other hand a \emph{bookings} is time elapsing two parking where the car is not tracked by the system. The intuition is to track the availability of each car on the car sharing platform, and rebuild the historic parking and booking periods over time: when a customer books a car, the latter ``disappears'' from the system. The framework records this event, with the initial time and position of a new booking. When the customer ends the booking, the car ``reappears'' in the system. The software records this event, with the final time and position of the booking. For the same car, a new parking period starts.

Harvested data is unstructured, and may grow large. Thus I leverage on \textit{MongoDB}, a NoSQL document-based database. A MongoDB database includes a set collections, i.e., groups of documents. Each document is a set of key-value pairs, organized in a JSON structure. The schema-less structure of MongoDB fits well in this work, because it can handle in the same collection documents defined with different key-value pairs. I decide to rely on such a system as I can easily manage the different field structures of providers, car2go and Enjoy in this use case. In addition, MongoDB offers a great integration with Python through the \textit{pymongo} module.

@@

\begin{table}
	\setlength{\tabcolsep}{2.3pt}
	\centering
	\caption{Overview of car2go's data}
	\begin{tabular}{lcccc}
		\hline
		City &  City Size [$km^2$]\footnote{} & Population \footnotemark[\value{footnote}] &  Avg. Fleat & Bookings\\ 
		\hline
		\hline
		Columbus			  & 583 & 892k & 187 & 186k \\
		Florence 	  			& 102 & 372k & 220 & 333k \\
		Denver 		  			 & 401 & 727k & 312 & 348k \\
		Austin 		   			  & 704 & 964k & 315 & 377k \\
		Frankfurt 			   & 248 & 701k & 242 & 505k \\
		Toronto 				 & 630 & 3120k & 400 & 536k \\
		Amsterdam 			& 219 & 854k & 314 & 573k \\
		Montreal 				& 431 & 1704k & 429 & 606k \\
		New York City(Manhattan) 	   & 59 & 1629 & 500 & 739k \\
		Turin 					    & 130 & 874k & 396 & 868k \\
		Munich 					 & 310 & 1464k & 478 & 916k \\
		Washington DC 	  & 177 & 705k & 563 & 919k \\
		Stuttgart 				 & 207 & 632k & 486 & 1001k \\
		Seattle 				   & 217 & 744k & 710 & 1134k \\
		Calgary 				  & 825 & 1239k & 552 & 1176k \\
		Rome 					  & 1287 & 2837k & 582 & 1240k \\
		Rheinland 			   &  - & 1688k & 648 & 1421k \\
		Vienna 					  & 414 & 1915k & 688 & 1702k \\
		Madrid 					 & 604 & 3233k & 424 & 2092k \\
		Milan 					   & 181 & 1396k & 776 & 2223k \\
		Hamburg 			  & 755 & 1833k & 812 & 2561k \\
		Vancouver 			 & 115 & 631k & 977 & 2701k \\
		Berlin 					  & 891 & 3769k & 1009 & 3091k \\
		%\\ \hline
		\hline
		\label{tab:2_3_datasummary}
	\end{tabular}
\end{table}

\begin{table}
	\setlength{\tabcolsep}{2.3pt}
	\centering
	\caption{Overview of Enjoy's data}
	\begin{tabular}{lcccc}
		\hline
		City &  City Size [$km^2$]\footnotemark[\value{footnote}] & Population \footnotemark[\value{footnote}] &  Avg. Fleat & Bookings\\ 
		\hline
		\hline
		Bologna			         &140    & 390k & 72 & 47k \\
		Catania			          &182    & 311k & 78 & 205k \\	
		Florence 	  			 & 102  & 372k & 82 & 279k \\
		Turin 					   & 130   & 874k & 251 & 983k \\
		Rome 					  & 1287 & 2837k & 601 & 2031k \\
		Milan 					   & 181   & 1396k & 755 & 3107k \\

@@

The first stage explains the data structures and how I snapshots the system getting all the car ready for a ride.

>>>>

\section{Data Normalization and Integration}
\label{sec:2_4_data_normalization}
In this second module I illustrates how \tool processes and consolidate each snapshot to obtain \emph{parking} and \emph{bookings} periods for each car. A \emph{parking} is time where the car is available for a user ride. On the other hand a \emph{bookings} is time between two parkings where the car is not tracked by the system. The intuition is to track the availability of each car on the car sharing platform, and rebuild the historic parking and booking periods over time: when a customer books a car, the latter ``disappears'' from the system. \tool records this event, with the initial time and position of a new booking. When the customer ends the booking, the car ``reappears'' in the system. \tool records this event, deriving the final time and position of the booking. For the same car, a new parking period starts.

Harvested data is unstructured, and may grow large. Thus I leverage \textit{MongoDB}, a NoSQL document-based database. A MongoDB database includes a set collections, i.e., groups of documents. Each document is a set of key-value pairs, organized in a JSON structure. The schema-less structure of MongoDB fits well in this work, because it can handle in the same collection documents defined with different key-value pairs. I decide to rely on such a system as I can easily manage the different field structures of providers, car2go and Enjoy in this use case. In addition, MongoDB offers a great integration with Python through the \textit{pymongo} module.

@@

\begin{table}
	\setlength{\tabcolsep}{2.3pt}
	\centering
	\caption{Overview of car2go's data sorted by number of bookings}
	\begin{tabular}{lcccc}
		\hline
		City &  City Size [$km^2$]\footnote{} & Population \footnotemark[\value{footnote}] &  Avg. Fleat & Bookings\\ 
		\hline
		\hline
		Columbus			  & 583 & 892k & 187 & 186k \\
		Florence 	  			& 102 & 372k & 220 & 333k \\
		Denver 		  			 & 401 & 727k & 312 & 348k \\
		Austin 		   			  & 704 & 964k & 315 & 377k \\
		Frankfurt 			   & 248 & 701k & 242 & 505k \\
		Toronto 				 & 630 & 3120k & 400 & 536k \\
		Amsterdam 			& 219 & 854k & 314 & 573k \\
		Montreal 				& 431 & 1704k & 429 & 606k \\
		New York City(Manhattan) 	   & 59 & 1629 & 500 & 739k \\
		Turin 					    & 130 & 874k & 396 & 868k \\
		Munich 					 & 310 & 1464k & 478 & 916k \\
		Washington DC 	  & 177 & 705k & 563 & 919k \\
		Stuttgart 				 & 207 & 632k & 486 & 1001k \\
		Seattle 				   & 217 & 744k & 710 & 1134k \\
		Calgary 				  & 825 & 1239k & 552 & 1176k \\
		Rome 					  & 1287 & 2837k & 582 & 1240k \\
		Rheinland 			   &  - & 1688k & 648 & 1421k \\
		Vienna 					  & 414 & 1915k & 688 & 1702k \\
		Madrid 					 & 604 & 3233k & 424 & 2092k \\
		Milan 					   & 181 & 1396k & 776 & 2223k \\
		Hamburg 			  & 755 & 1833k & 812 & 2561k \\
		Vancouver 			 & 115 & 631k & 977 & 2701k \\
		Berlin 					  & 891 & 3769k & 1009 & 3091k \\
		%\\ \hline
		\hline
		\label{tab:2_3_datasummary}
	\end{tabular}
\end{table}

\begin{table}
	\setlength{\tabcolsep}{2.3pt}
	\centering
	\caption{Overview of Enjoy's data sorted by number of bookings}
	\begin{tabular}{lcccc}
		\hline
		City &  City Size [$km^2$]\footnotemark[\value{footnote}] & Population \footnotemark[\value{footnote}] &  Avg. Fleat & Bookings\\ 
		\hline
		\hline
		Bologna			         &140    & 390k & 72 & 47k \\
		Catania			          &182    & 311k & 78 & 205k \\	
		Florence 	  			 & 102  & 372k & 82 & 279k \\
		Turin 					   & 130   & 874k & 251 & 983k \\
		Rome 					  & 1287 & 2837k & 601 & 2031k \\
		Milan 					   & 181   & 1396k & 755 & 3107k \\

@@

The first stage explains the data structures and how I take snapshots the system getting all the car ready for a ride.

### 2020-12-04 Chapter3/3_5_rides_characterization.tex
added appendix

Figure~\ref{fig:3_5_cdf_corrective_factors} reports the ECDF of the ratio between the Google Distance and the straight distance. The plot shows how, in the 95\% of rides, the Google distance is about twice time bigger than the straight distance. However, in average it is possible to observe a multiplicative factor of 1.5 for and enjoy. Anyway, in the following chapters and in particular in chapters \ref{chap:5_simulator}, \ref{chap:6_4cities} and \ref{chap:7_cs_optimization} I used the median value, of 1,4. Notice that the bigger values in car2go are due to a bigger operative area. The table \ref{tab:3_5_corrective_factor} resumes the main ECDF values.

\begin{table}[h!]
	\centering
	\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}

		\hline
		\textbf{Provider} & \textbf{Mean} & \textbf{Std} & \textbf{Min} & \textbf{25\%} & \textbf{50\%} & \textbf{75\%} & \textbf{90\%} & \textbf{95\%} & \textbf{99\%} \\ \hline
		car2go            & 1.5           & 0.54         & 0.53         & 1.27          & 1.40          & 1.59          & 1.90          & 2.21          & 3.47          \\ \hline
		enjoy             & 1.48          & 0.52         & 0.87         & 1.26          & 1.37          & 1.54          & 1.81          & 2.12          & 3.38          \\ \hline
	\end{tabular}
	\label{tab:3_5_corrective_factor}	
\end{table}

>>>>

Figure~\ref{fig:3_5_cdf_corrective_factors} reports the ECDF of the ratio between the Google Distance and the straight distance. The plot shows how, in the 95\% of rides, the Google distance is about twice time bigger than the straight distance. However, in average it is possible to observe a multiplicative factor of 1.5 for and enjoy. Anyway, in the following chapters and in particular in chapters \ref{chap:5_simulator}, \ref{chap:6_4cities} and \ref{chap:7_cs_optimization} I used the median value, of 1,4. Notice that the bigger values in car2go are due to a bigger operative area. Table \ref{tab:3_5_corrective_factor} resumes the main ECDF values.

\begin{table}[h!]
	\centering
	\caption{Statistics on corrective factor for car2go and Enjoy}
	\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}

		\hline
		\textbf{Provider} & \textbf{Mean} & \textbf{Std} & \textbf{Min} & \textbf{25\%} & \textbf{50\%} & \textbf{75\%} & \textbf{90\%} & \textbf{95\%} & \textbf{99\%} \\ \hline
		car2go            & 1.5           & 0.54         & 0.53         & 1.27          & 1.40          & 1.59          & 1.90          & 2.21          & 3.47          \\ \hline
		Enjoy             & 1.48          & 0.52         & 0.87         & 1.26          & 1.37          & 1.54          & 1.81          & 2.12          & 3.38          \\ \hline
%		\caption{Statistics about corrective factor for car2go and Enjoy}}

	\end{tabular}
	\label{tab:3_5_corrective_factor}
\end{table}

### 2020-12-04 Chapter4/4_0_abstract.tex
added appendix

This chapter refers mostly the paper \textit{Characterizing client usage patterns and service demand for car-sharing systems} \cite{VancouverCS}, published on the Journal, available online since October 11, 2019. My contribution is mainly focused in all the aspects about car2go analyses.

>>>>

This chapter refers mostly the paper \textit{Characterizing client usage patterns and service demand for car-sharing systems} \cite{VancouverCS}, published on the Journal Information Systems, available online since October 11, 2019. My contribution is mainly focused in all the aspects about car2go analyses.

### 2020-12-04 Chapter5/5_1_introduction.tex
added appendix

\section{Introduction}
In the previous chapters I widely described the dataset in I collected and characterized. However the collected data refers \emph{only} to internal combustion engine cars. Recalling that the main research question of my thesis is: \emph{It is possible design an electric Free Floating Car Sharing System?}, I need a tool able to replicate the FFCS customers' habits in a customizable model of electric FFCS system.

In this chapter I describe the core-tool of my research: an event-driven trace-based electric FFCS simulator. In a nutshell, this software I developed is able to (i)  extract users pattern from data collected in chapter \ref{chap:2_dataset}; (ii) replicate the users patterns in a electrified scenario parametrized by the client; (iii) return several performance metrics describing the inputted electric FFCS setup.

@@

Those metrics are heavily influenced by the some environmental parameters like the number and the distribution of charging station. For this reason I proposed three placement strategies related to users' driving patterns.

>>>>

\section{Introduction}
In the previous chapters I widely described the dataset I collected and characterized. However the collected data refers \emph{only} to internal combustion engine cars. Recalling that the main research question of my thesis is: \emph{It is possible design an electric Free Floating Car Sharing System?} I need a tool able to replicate the FFCS customers' habits in a customizable model of electric FFCS system.

In this chapter I describe the core-tool of my research: an event-driven trace-based electric FFCS simulator. In a nutshell, this software is able to (i)  extract users pattern from data collected in chapter \ref{chap:2_dataset}; (ii) replicate the users patterns in a electrified scenario parametrized by the client; (iii) return several performance metrics describing the inputted electric FFCS setup.

@@

Those metrics are heavily influenced by the some environmental parameters like the number and the distribution of charging station. For this reason I propose three placement strategies related to users' driving patterns.

### 2020-12-04 Chapter5/5_5_simulation_scenario.tex
added appendix

I used this simulator to study the impact on the number of zones that are equipped with charging stations $N$, and the number of poles $k$ of each charging station.

>>>>

I use this simulator to study the impact on the number of zones that are equipped with charging stations $N$, and the number of poles $k$ of each charging station.

### 2020-12-04 thesis.tex
added appendix

\summary%\sommario
%This is where you write your abstract \dots\ (Maximum 4000 characters, i.e. maximum two pages in normal sized font, typeset with the thesis layout).
%The abstract environment is also available, but  \texttt{\string\summary} is preferred because it generates an un-numbered chapter. The abstract environment is more suitable for articles and two column typesetting without a separate title page.
Nowadays, the increase of traffic congestions, land consume and pollution emission due to a private car ownership opened the way to share mobility. One of the most spread implementation of shared mobility is Free Floating Car Sharing (FFCS). It is a car rental model where the users can pick and release the car everywhere within an operative area. The customers can reserve (and release) the vehicle through web-based application. With just a simple tap, the users can unlock and lock the smart vehicles. Usually the provider bills the users only for the time spend driving, with time-minute based fares. All the other costs like petrol, insurance and maintenance are in charge to the provider.

@@

Given the recent increase market of the electric vehicles and all the pros those vehicles carry, replace FFCS fleet with electric powered cars may still improve the quality of life in urban centres. The setup and management of an electric FFCS requires ingenuity, in order to minimize the users discomfort due the plugging procedures .

In my thesis, I presented a methodology to address, in different cases of studies, all the challenges related to the conversion of combustion engine cars to electric vehicles in FFCS. In particular, the main drivers focus on a profitable and technically sustainable system setup able to guarantee a flexible and appealing mobility service to an increasing customer audience.

In the first part of my thesis, I described the software I developed in order to scrape from the web real combustion engine FFCS, from two providers: car2go and Enjoy. The data collection car2go data collection lasted from \mc{december 2016 to january 2018, collecting more than 27 million users' bookings spread in 23 cities} while the Enjoy data collection phase started in May 2017 and lasted until June 2019, recording about 6 million bookings in 6 cities.

Then, I characterized both datasets in Turin, one of the cities in which both FFCS providers work. I defined outlier detection filtering criteria and I extracted geo-temporal users' travel patterns.

After that, I compared the car2go customer's pattern with the one-way and two-way car-sharing system. The results showed how users prefer more flexible services like FFCS or one-way car sharing.

Once the data are consolidated, I developed (i) a methodology to assign a charging station to a city by looking at users' pattern; (ii) system policies to manage the fleet when the vehicle SoC may not guarantee a trip; (iii) an event-based trace-driven simulator able to replicate the recorded trips in an electrified scenario evaluating the feasibility of each configuration.

Then, via accurate simulation in Berlin, Milan, Turin and Vancouver, I studied different electric FFCS setup. By placing the charging station in the most frequented areas, incentive the users to plug the car when the battery state of charge is below a safety threshold and balancing the spread of poles, it is possible to obtain a sustainable system covering with chaging station only the 8-10~\% of zones

To still reduce the number of charging stations to have a sustainable electric FFCS, I compared several optimization algorithms. The results showed how a Genetic Algorithm is able to find a better solution able to shrink the minimum amount of resources to sustain the same mobility demand.

After that, I moved my attention to the users' rentals predictability. The main goal is to understand how different open-data sources could impact on the recorded FFCS users' rental. Initially, I compared several time-series forecasts to predict the users' demand in the short and medium-term. The Random Forrest regression produced better accuracy and results interpretability. Then I correlated the socio-economics features characterizing each Vancouver neighborhood to FFCS demand and again, the Random Forrest regression outperforms other algorithms

Finally, I questioned the system scalability figuring out several scenarios having increasing demand. I used a model able to synthesize users' demand by looking only at the geo-spatial users' rentals. By varying the electric FFCS setup and simulating the new scenario I pointed out how a linear increase in the demand intensity leads to a fleet sub-linear increase. Finally, I projected those considerations in euros proofing how electric FFCS has rooms for economic growth.

\ringraziamenti
tba

@@

\include{Chapter1/1_introduction}
\include{Chapter2/2_datacollection}
\include{Chapter3/3_datacharact}
\include{Chapter4/4_cs_comparison}
\include{Chapter5/5_simulator}
\include{Chapter6/6_4cities}
\include{Chapter7/7_cs_optimization}
\include{Chapter8/8_prediction}
%\include{Chapter9/9_relocation}
\include{Chapter10/10_scalability}
\include{Chapter11/11_conclusion}
%% Numbered appendices remain in the main matter...
%\appendix
%\include{Appendix1/appendix1}
%\include{Appendix2/appendix2}

>>>>

\summary%\sommario
%This is where you write your abstract \dots\ (Maximum 4000 characters, i.e. maximum two pages in normal sized font, typeset with the thesis layout).
%The abstract environment is also available, but  \texttt{\string\summary} is preferred because it generates an un-numbered chapter. The abstract environment is more suitable for articles and two column typesetting without a separate title page.
Nowadays, the increase of traffic congestions, land consume and pollution emission due to a private car ownership opened the way to shared mobility. One of the most spread implementation of shared mobility is Free Floating Car Sharing (FFCS). It is a car rental model where the users can pick and release the car everywhere within an operative area. The customers can reserve (and return) the vehicle through web-based application. With just a simple tap, the users can unlock and lock the smart vehicle. Usually the provider bills the users only for the time spend driving, with time-minute based fares. All the other costs like petrol, insurance and maintenance are in charge to the provider.

@@

Given the recent market market of the electric vehicles and all the pros those vehicles carry, replace FFCS fleet with electric powered cars may still improve the quality of life in urban centres. The setup and management of an electric FFCS requires ingenuity, in order to minimize the users discomfort due the plugging procedures.

In my thesis, I present a methodology to address, in different cases of studies, all the challenges related to the conversion of combustion engine cars to electric vehicles in FFCS. In particular, the main drivers focus on a profitable and technically sustainable system setup able to guarantee a flexible and appealing mobility service to an increasing customer audience.

In the first part of my thesis, I describe the software I developed in order to scrape from the web real combustion engine FFCS, from two providers: car2go and Enjoy. The car2go data collection lasted from December 2016 to January 2018, collecting more than 27 million users' bookings spread in 23 cities while the Enjoy data collection phase started in May 2017 and lasted until June 2019, recording about 6 million bookings in 6 cities.

Then, I characterize both datasets in Turin, one of the cities in which both FFCS providers work. I detect the outliers filtering them out from the dataset and I extract geo-temporal users' travel patterns.

After that, I compare the car2go customer's pattern with the one-way and two-way car-sharing system. The results show how users prefer more flexible services like FFCS or one-way car sharing.

Once the data are consolidated, I develop (i) a methodology to place a charging station to a city by looking at users' pattern; (ii) system policies to manage the fleet when the vehicle state of charge may not guarantee a trip; (iii) via an event-based trace-driven simulator able to replicate the recorded trips in an electrified scenario evaluating the feasibility of each configuration.

Via accurate simulation in Berlin, Milan, Turin and Vancouver, I study different electric FFCS setup. By placing the charging station in the most frequented areas, by offering incentive the users to plug the car when the battery state of charge is below a safety threshold and balancing the spread of poles, it is possible to obtain a sustainable system covering with chaging station only the 8-10~\% of zones

To still reduce the number of charging stations to have a sustainable electric FFCS, I compare several optimization algorithms. The results show how a Genetic Algorithm is able to find a better solution able to shrink the minimum amount of resources to sustain the same mobility demand.

After that, I move my attention to the users' rentals predictability. The main goal is to understand how different open-data sources could impact on the recorded FFCS users' rental. Initially, I compare several time-series forecasts to predict the users' demand in the short and medium-term. The Random Forest regression produce better accuracy and results interpretability. Then I correlate the socio-economics features characterizing each city neighbourhood to FFCS demand and again, the Random Forest regression outperforms other algorithms

Finally, I question the system scalability figuring out several scenarios having increasing demand. I use a model able to synthesize users' demand by looking only at the geo-spatial users' rentals. By varying the electric FFCS setup and simulating the new scenario I point out how a linear increase in the demand intensity requires a fleet sub-linear increase. Finally, I project those considerations in euros proofing how electric FFCS has rooms for economic growth.

%\ringraziamenti
%tba

@@

\include{Chapter1/1_introduction}
\include{Chapter2/2_datacollection}
\include{Chapter3/3_datacharact}
\include{Chapter4/4_cs_comparison}
\include{Chapter5/5_simulator}
\include{Chapter6/6_4cities}
\include{Chapter7/7_cs_optimization}
\include{Chapter8/8_prediction}
%\include{Chapter9/9_relocation}
\include{Chapter10/10_scalability}
\include{Chapter11/11_conclusion}
%% Numbered appendices remain in the main matter...
\appendix
\include{Appendix1/appendix1}
%\include{Appendix2/appendix2}
